---
title: "Heatmap of time (hour) by trip destination"
date: "11/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)

```

```{r}
heatmap <-
  read_csv("./data/pilot.csv")

heatmap %>% 
  filter(do_boro != "Unknown") %>% 
  select (do_boro, pu_hour) %>% 
  mutate(
    do_boro = fct_infreq(do_boro)
  ) %>% 
  group_by(do_boro, pu_hour) %>% 
  count() %>% 
  group_by(do_boro) %>% 
  mutate( 
    total = sum(n)
    ) %>% 
   mutate(percent = round((n/total)*100,1)) %>% 
  arrange(do_boro, pu_hour) %>% 
    plot_ly(
    x = ~pu_hour, y = ~do_boro,
    height = 400, width = 1400) %>%
  add_heatmap(
    z = ~percent,
    text = ~paste(
      "Trip to: ", do_boro, "<br>Time: ", pu_hour, ": 00",
      "<br>Percent: ", percent, "%"),
    hoverinfo = "text", showscale = TRUE,
    opacity = 0.90)



```




Each square in the heatmap represents a unique combination of destination boroughs and 24 hours in a day, and the color represents the percent of trip amount out of the total trip amount in that  borough,  with brighter color represents higher percent.
   
We can see that in boroughs like Staten Island, Bronx, or Brroklyn, there were more trips to these places in night or late night (8pm - 1am) than in the day time. Trips to Queens were relatively evenly spread out across hours in a day and trips to Manhattan were more likely to be in the day  rather than in the night. Trips to Newark had a clear peak in the afternoon hours.